dispRity manual

### 2017-06-16 doc/dispRity-manual.Rmd
Updated manual

@@

An alternative way to display the calculated disparity is to plot the results using the S3 method `plot.dispRity`.
This function intakes the same options as `summary.dispRity` along side with various graphical options described in the function manual (see `?plot.dispRity`).

@@

```{r, fig.width=6, fig.height=6, eval=FALSE}
## Graphical parameters
op <- par(mfrow = c(2,2), bty = "n")
## Plotting continuous disparity results
plot(disp_time_binsBS, type = "continuous")
## Plotting discrete disparity results
plot(disp_crown_stemBS, type = "box")
## Plotting the disparity using lines for the rarefaction level of 20 elements only
plot(disp_crown_stemBS, type = "line", rarefaction = 20)
## Plotting the disparity using polygons while also displaying the number of elements
plot(disp_crown_stemBS, type = "polygon", elements = TRUE)
```

> The `type` argument is not mandatory for the two first plots since both `continuous` and `box` are the default options for respectively continuous and discrete data.

@@

```{r, fig.width=6, fig.height=6, eval=FALSE}
## Graphical options
par(bty = "n")
## Plotting the results with some classic options from plot
plot(disp_time_slicesBS, col = c("blue", "orange", "green"), ylab = c("Some measurement"),
    xlab = "Some other measurement", main = "Many options...", ylim = c(0,2.1))
## Adding a legend
legend("topleft", legend = c("Central tendency", "Confidence interval 1", "Confidence interval 2"),
    col = c("blue", "orange", "green"), pch = 19)
```

@@

```{r, fig.width=6, fig.height=6, eval=FALSE}
## Graphical options
par(bty = "n")
## Plotting the results with some plot.dispRity arguments
plot(disp_time_slicesBS, quantile = c(seq(from=10, to=100, by=10)), cent.tend = sd, type = "c",
    elements = TRUE, col = c("black", rainbow(10)), ylab = c("Disparity", "Diversity"),
    xlab = "Time (in in units from past to present)", time.subsamples = FALSE,
    observed = TRUE, main = "Many options...")
```

@@

```{r, fig.width=6, fig.height=6, eval=FALSE}
## Graphical options
par(bty = "n")
## Plotting the continuous disparity with a fixed y axis
plot(disp_time_slicesBS, ylim = c(0.5,2))
## Adding the discrete data using the same fixed y axis to have the results at the same scale
plot(disp_time_binsBS, type = "line", ylim=c(0.5,2), xlab="", ylab="", add = TRUE)
```

@@

```{r, fig.width=6, fig.height=6, eval=FALSE}
## Graphical options
par(bty = "n")
## Plotting the rarefaction curves
plot(disp_crown_stemBS, rarefaction = TRUE)
```

>>>>

## Plotting the results
plot(disparity_data, type = "continuous")

@@

## Summarising dispRity data (plots)

Because of its architecture, printing `dispRity` objects only summarise their content but does not print the disparity value measured or associated analysis.
<!-- link to dispRity objects -->
To actually summary what's in a dispRity object, one can either use the `summary` function for visualising the data in a table or `plot` to have a graphic representation of the results.

### Summarising `dispRity` data
This function is a S3 function (`summary.dispRity`) allowing to summarise the content of `dispRity` objects that contain disparity calculations.

```{r}
## Example data from previous sections
crown_stem <- custom.subsamples(BeckLee_mat50,
    group = list("crown" = c(16, 19:41, 45:50), "stem" = c(1:15, 17:18, 42:44)))
## Bootstrapping and rarefying these groups
boot_crown_stem <- boot.matrix(crown_stem, bootstraps = 200, rarefaction = TRUE)
## Calculate disparity
disparity_crown_stem <- dispRity(boot_crown_stem, metric = c(sum, variances))

## Creating time slices subsamples
time_slices <- time.subsamples(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "gradual", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)
## Bootstrapping on the time binning/slicing subsamples
boot_time_slices <- boot.matrix(time_slices, bootstraps = 1000)
## Calculate disparity
disparity_time_slices <- dispRity(boot_time_slices, metric = c(sum, variances))
```

This objects are easy to summarise as follow:

```{r}
## Default summary
summary(disparity_time_slices)
```

Information about the number of elements per subsamples as well as the observed (i.e. non-bootstrapped) disparity is also calculated.
This is specifically handy when rarefying the data for example

```{r}
head(summary(disparity_crown_stem))
```

The summary functions can also take various options such as the `quantile` values for the confidence intervals levels (by default, the 50 and 95 quantiles are calculated); the `cent.tend` for the  central tendency to use for summarising the results (default is `median`); the `rounding` option corresponding to the number of decimals to print (default is `2`) and the `recall` option for printing the call of the `dispRity` object as well (default is `FALSE`)
These options can easily be changed from the defaults as follows:

```{r}
## Same but using the 88 quantile and the standard deviation as the summary. 
summary(disparity_time_slices, quantile = 88, cent.tend = sd)

## Printing the details of the object and rounding the values to the 5th decimal
summary(disparity_time_slices, recall = TRUE, rounding = 5)
```

Information about the number of elements per subsamples as well as the observed (i.e. non-bootstrapped) disparity is also calculated.
This is specifically handy when rarefying the data for example

Note that the summary table is a `data.frame`, hence it is easy to modify as any table using `dplyr`.
You can also export it in `csv` format using `write.csv` or even directly export into `LaTeX` format using the following;

```{r, eval = FALSE}
## Loading the xtable package
require(xtable)
## Converting the table in LaTeX
xtable(summary(disparity_time_slices))
```

### Plotting `dispRity` data
An alternative (and more fun!) way to display the calculated disparity is to plot the results using the S3 method `plot.dispRity`.
This function intakes the same options as `summary.dispRity` along side with various graphical options described in the function manual (see `?plot.dispRity`).

@@

```{r, fig.width=8, fig.height=8}
## Graphical parameters
op <- par(mfrow = c(2,2), bty = "n")

## Plotting continuous disparity results
plot(disparity_time_slices, type = "continuous")

## Plotting discrete disparity results
plot(disparity_crown_stem, type = "box")

## Same but using lines for the rarefaction level of 20 elements only
plot(disparity_crown_stem, type = "line", rarefaction = 20)

## Same but using polygons while also displaying the number of elements
plot(disparity_crown_stem, type = "polygon", elements = TRUE)

## Resetting graphical parameters
par(op)
```

@@

```{r, fig.width=8, fig.height=8}
## Graphical options
par(bty = "n")

## Plotting the results with some classic options from plot
plot(disparity_time_slices, col = c("blue", "orange", "green"),
    ylab = c("Some measurement"), xlab = "Some other measurement",
    main = "Many options...", ylim = c(5, 0), xlim = c(4, 0))

## Adding a legend
legend("topleft", legend = c("Central tendency",
                             "Confidence interval 1",
                            "Confidence interval 2"),
     col = c("blue", "orange", "green"), pch = 19)

## Resetting graphical parameters
par(op)
```

@@

```{r, fig.width=8, fig.height=8}
## Graphical options
par(bty = "n")

## Plotting the results with some plot.dispRity arguments
plot(disparity_time_slices, quantile = c(seq(from=10, to=100, by=10)),
    cent.tend = sd, type = "c", elements = TRUE, col = c("black", rainbow(10)),
    ylab = c("Disparity", "Diversity"), time.subsamples = FALSE,
    xlab = "Time (in in units from past to present)", observed = TRUE,
    main = "Many more options...")

## Resetting graphical parameters
par(op)
```

@@

```{r, fig.width=8, fig.height=8}
## Graphical options
par(bty = "n")

## Plotting the continuous disparity with a fixed y axis
plot(disparity_time_slices, ylim = c(1, 4))
## Adding the discrete data
plot(disparity_crown_stem, type = "line", ylim = c(1, 4), xlab = "", ylab = "",
    add = TRUE)

## Resetting graphical parameters
par(op)
```

<!-- Change this example to add time bins! -->

@@

```{r, fig.width=8, fig.height=8}
## Graphical options
par(bty = "n")

## Plotting the rarefaction curves
plot(disparity_crown_stem, rarefaction = TRUE)

## Resetting graphical parameters
par(op)
```
